\chapter{\texorpdfstring{Vulnerability 5.06a: Function Pointer\\Type Incompatibility}{Vulnerability 5.06a: Function Pointer Type Incompatibility}}

\section{Executive Summary}

Vulnerability 5.06a is an ABI (Application Binary Interface) mismatch where a function pointer is declared with one signature (taking no arguments) but assigned to and called as a function with a completely different signature (expecting arguments). This violates ISO-IEC Rule 5.06 and represents a \textbf{verifier bypass} rather than a directly exploitable memory corruption.

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=ISO Rule Violated,fonttitle=\bfseries]
ISO-IEC Rule 5.06: ``Do not declare pointers to functions with a type that differs from the function definition''
\end{tcolorbox}

\begin{tcolorbox}[colback=yellow!5!white,colframe=orange!75!black,title=Severity,fonttitle=\bfseries]
\textbf{MEDIUM:} Verifier bypass confirmed. No direct memory corruption in this context, but demonstrates systematic ABI enforcement gap in the verifier.
\end{tcolorbox}

\subsection{Why This Vulnerability Matters}

\begin{enumerate}[label=\arabic*.]
  \item \textbf{Verifier Gap}: The eBPF verifier does not validate function pointer type compatibility
  \item \textbf{ABI Violation}: Function arguments are passed in fixed registers (R1-R5), but type mismatches mean the caller and callee disagree on what those values represent
  \item \textbf{Systemic Risk}: While limited in this specific code context, similar patterns in pointer arithmetic or array indexing could become dangerous
\end{enumerate}

\section{The Vulnerable Code Pattern}

\subsection{Vulnerable Code (Simplified)}

\vspace{0.3cm}
\noindent
\fcolorbox{black!30}{gray!15}{%
  \begin{minipage}[c]{\dimexpr\textwidth-2\fboxsep-2\fboxrule}
    {\small\ttfamily
    /* Forward declaration with wrong signature */\\
    \_\_u16 (*fp\_wrong\_type)();  /* Takes no arguments */\\
    \\
    /* Assignment to a function expecting different signature */\\
    fp\_wrong\_type = csum\_fold;  /* csum\_fold expects \_\_u32 */\\
    \\
    /* Call with incompatible arguments */\\
    \_\_u16 csum\_result = fp\_wrong\_type(0x12345678, 0x9abcdef0);\\
    bpf\_printk("[argcomp]: 5.6a - csum: \%u", csum\_result);
    }
  \end{minipage}%
}
\vspace{0.3cm}

\subsection{Why This Is Vulnerable}

\begin{enumerate}[label=\arabic*.]
  \item \textbf{Declared Type}: The pointer declares `\_\_u16 (*)(void)` takes no arguments
  \item \textbf{Actual Function}: But it's assigned to `csum\_fold`, which expects one `\_\_u32` argument
  \item \textbf{Called With}: Two arguments (`0x12345678`, `0x9abcdef0`) are passed
  \item \textbf{Verifier Gap}: The verifier focuses on memory safety, not ABI (Application Binary Interface) correctness. It doesn't cross-check the declared type against the actual function signature
\end{enumerate}

\subsection{What Should Prevent This}

The verifier should:
\begin{itemize}
  \item Track function pointer types throughout their lifetime
  \item Validate that calls through pointers match the declared prototype
  \item Reject assignments that violate type contracts
\end{itemize}
\noindent
Currently, \textbf{it does not perform this full validation}, allowing 5.06a to pass.

\section{Multi-Stage Analysis: Evidence Chain}

\subsection{Stage 1: Source Code (Patch Location)}

The vulnerability is introduced via a Git patch: 

\begin{tcolorbox}[colback=gray!10, colframe=gray!10, arc=2pt, boxrule=0pt]
\small\ttfamily
XDPs/xdp\_synproxy/patches/5\_06a\_argcomp/0001-feat-5.06a-function\_pointer-type-incompatibility.patch
\end{tcolorbox}

\subsection{Stage 2: Bytecode (llvm-objdump)}

Using \texttt{llvm-objdump -S xdp\_synproxy\_kern.o}, the compiled eBPF bytecode preserves the vulnerable instruction sequence:

\vspace{0.3cm}
\noindent
\fcolorbox{black!30}{gray!15}{%
  \begin{minipage}[c]{\dimexpr\textwidth-2\fboxsep-2\fboxrule}
    {\small\ttfamily
    ;~ bpf\_printk("[argcomp]: 5.6a - csum: \%u", csum\_result);\\
    56: (b7) r3 = 0x9753  \phantom{xx}  \# Load constant 38739\\
    57: (85) call 0x6     \phantom{xx}  \# Call bpf\_trace\_printk
    }
  \end{minipage}%
}
\vspace{0.3cm}

\noindent
\textbf{Finding}: The constant `0x9753` (38739) appears in bytecode, proving the vulnerable code path was \textbf{not optimized away} by the compiler.

\subsection{Stage 3: Verifier (bpftool prog load)}

Command:
\begin{tcolorbox}[colback=gray!10, colframe=gray!10, arc=2pt, boxrule=0pt]
\small\ttfamily
\texttt{sudo bpftool prog load xdp\_synproxy\_kern.bpf.o /sys/fs/bpf/test\_synproxy type xdp -d}
\end{tcolorbox}

\noindent
\textbf{Result}: The verifier \textbf{accepts the program without error}. No rejection, no warning.

\noindent
\textbf{Finding}: The verifier does not validate function pointer type compatibility. The program passes verification despite the ABI (Application Binary Interface) mismatch.

\subsection{Stage 4: Xlated Bytecode (bpftool prog dump xlated)}

The verifier-processed bytecode:

\vspace{0.3cm}
\noindent
\fcolorbox{black!30}{gray!15}{%
  \begin{minipage}[c]{\dimexpr\textwidth-2\fboxsep-2\fboxrule}
    {\small\ttfamily
    ;~ bpf\_printk("[argcomp]: 5.6a - csum: \%u", csum\_result);\\
    56: (b7) r3 = 38739        \# Constant still present\\
    57: (85) call bpf\_trace\_printk\#-115712
    }
  \end{minipage}%
}
\vspace{0.3cm}

\noindent
\textbf{Finding}: The xlated bytecode is nearly identical to the compiled bytecode. The constant and function call \textbf{survive verification unchanged}. No guards or safety checks inserted.

\subsection{Stage 5: Runtime Execution (trace\_pipe)}

Trigger the vulnerability by sending TCP SYN packets:
\begin{tcolorbox}[colback=gray!10, colframe=gray!10, arc=2pt, boxrule=0pt]
\small\ttfamily
python3 src/exploits/5\_06a\_argcomp/exploit\_506a.py 
\end{tcolorbox}

\noindent
Kernel trace output:

\vspace{0.3cm}
\noindent
\fcolorbox{black!30}{gray!15}{%
  \begin{minipage}[c][0.8cm][c]{\dimexpr\textwidth-2\fboxsep-2\fboxrule}
    \texttt{<idle>-0 [000] ..s2. 348.737756: bpf\_trace\_printk: [argcomp]: 5.6a - csum: 38739}
  \end{minipage}%
}
\vspace{0.3cm}

\noindent
\textbf{Finding}: The constant value `38739` appears in kernel traces, linking all previous stages together. The vulnerable code executes at runtime.

\subsection{Summary: Evidence Chain}

\begin{center}
\begin{tabular}{|l|l|}
\hline
\textbf{Stage} & \textbf{Finding} \\
\hline
Bytecode & Constant \texttt{0x9753} present (not optimized away) \\
Verifier & No rejection; program accepted \\
Xlated & Constant and call survive verification \\
Runtime & Trace shows constant `38739` \\
\hline
\end{tabular}
\end{center}

\noindent
\textbf{Conclusion}: The vulnerability is \textbf{real, complete, and executes at runtime}.

\section{Exploitation Analysis}

In the current proof-of-concept, the mismatched call passes hardcoded values that are not dereferenced as pointers, so no memory corruption occurs. The function call completes, and the XDP handler continues normally. However, if similar ABI (Application Binary Interface) mismatches appeared in code that used the mismatched arguments for pointer arithmetic, array indexing, or as size parameters, they could produce invalid memory addresses, enable out-of-bounds access, or create buffer overflow conditions with false bounds. In those contexts, 5.06a would become \textbf{directly exploitable}.

\section{Artifact Evidence}

All analysis logs are available in: \texttt{src/exploits/5\_06a\_argcomp/logs/}

\begin{itemize}
  \item \texttt{5\_06a\_bytecode\_analysis.txt}, Full llvm-objdump output
  \item \texttt{5\_06a\_verifier\_log.txt}, Verifier output (bpftool prog load -d)
  \item \texttt{5\_06a\_xlated\_dump.txt}, Xlated bytecode (verifier-processed)
\end{itemize}

\noindent
PoC script: \texttt{src/exploits/5\_06a\_argcomp/exploit\_506a.py}
