\chapter{Vulnerability 5.06b: Wrong Number of Arguments}

\section{Executive Summary}

Vulnerability 5.06b is an ABI (Application Binary Interface) mismatch where a function is called with \textbf{more arguments than its signature declares}. This violates ISO-IEC Rule 5.06 and represents a \textbf{verifier bypass} rather than directly exploitable memory corruption in this context.

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=ISO Rule Violated,fonttitle=\bfseries]
ISO-IEC Rule 5.06: ``For a call to a function without a function prototype in scope, the number of arguments does not equal the number of parameters''
\end{tcolorbox}

\begin{tcolorbox}[colback=yellow!5!white,colframe=orange!75!black,title=Severity,fonttitle=\bfseries]
\textbf{MEDIUM:} Verifier bypass confirmed. Extra arguments are ignored by the callee, but register state is perturbed. Potential information disclosure vector.
\end{tcolorbox}

\subsection{Why This Vulnerability Matters}

\begin{enumerate}[label=\arabic*.]
  \item \textbf{Verifier Gap}: The eBPF verifier does not validate that the number of arguments matches the function signature
  \item \textbf{Register Misalignment}: Arguments are passed in fixed registers (R1-R5). Extra arguments leave register state inconsistent
  \item \textbf{Information Leakage}: If extra arguments contain sensitive kernel values, they may be disclosed to attackers
\end{enumerate}

\section{The Vulnerable Code Pattern}

\subsection{Vulnerable Code (Simplified)}

\vspace{0.3cm}
\noindent
\fcolorbox{black!30}{gray!15}{%
  \begin{minipage}[c]{\dimexpr\textwidth-2\fboxsep-2\fboxrule}
    {\small\ttfamily
    /* Forward declaration WITHOUT parameters */\\
    void network\_copy\_helper();\\
    \\
    /* Call with 3 arguments */\\
    \_\_u8 extra\_buf[ETH\_ALEN] = \{0\};\\
    network\_copy\_helper(hdr-\textgreater{}eth-\textgreater{}h\_source,\\
    \phantom{network\_copy\_helper(} hdr-\textgreater{}eth-\textgreater{}h\_dest,\\
    \phantom{network\_copy\_helper(} extra\_buf);\\
    \\
    /* Actual definition takes only 2 arguments */\\
    void network\_copy\_helper(char *dst, char *src) \{\\
      \_\_builtin\_memcpy(dst, src, ETH\_ALEN);\\
    \}
    }
  \end{minipage}%
}
\vspace{0.3cm}

\subsection{Why This Is Vulnerable}

\begin{enumerate}[label=\arabic*.]
  \item \textbf{Forward declaration}: Has \textbf{no parameter list} — `()` means unknown arity
  \item \textbf{Called with 3 arguments}: Passes values in R1, R2, R3
  \item \textbf{Function expects 2}: Only consumes R1 and R2
  \item \textbf{Verifier gap}: Doesn't check that argument count matches; only validates memory safety
  \item \textbf{Result}: R3 contains uninitialized/garbage register state
\end{enumerate}

\section{Multi-Stage Analysis: Evidence Chain}

\subsection{Stage 1: Source Code (Patch Location)}

The vulnerability is introduced via a Git patch: 
\begin{tcolorbox}[colback=gray!10, colframe=gray!10, arc=2pt, boxrule=0pt]
\small\ttfamily
XDPs/xdp\_synproxy/patches/5\_06b\_argcomp/0001-feat-5.06b-wrong-argument-count.patch
\end{tcolorbox}

\subsection{Stage 2: Bytecode (llvm-objdump)}

Using \texttt{llvm-objdump -S xdp\_synproxy\_kern.o}:

\vspace{0.3cm}
\noindent
\fcolorbox{black!30}{gray!15}{%
  \begin{minipage}[c]{\dimexpr\textwidth-2\fboxsep-2\fboxrule}
    {\small\ttfamily
    ;~ \_\_builtin\_memcpy(dst, src, ETH\_ALEN);\\
    61: (71) r1 = *(u8 *)(r4 +5)\\
    ;~ ... memcpy loop operations ...\\
    72: (73) *(u8 *)(r2 +6) = r1\\
    ;~ bpf\_printk("[argcomp]: 5.6b - network copy called");\\
    73: (18) r1 = 0x0 ll\\
    76: (85) call 0x6
    }
  \end{minipage}%
}
\vspace{0.3cm}

\noindent
\textbf{Finding}: The bytecode preserves the full call sequence including memory operations and the argument-count-mismatch signal. Not optimized away.

\subsection{Stage 3: Verifier (bpftool prog load)}

Command:
\begin{tcolorbox}[colback=gray!10, colframe=gray!10, arc=2pt, boxrule=0pt]
\small\ttfamily
sudo bpftool prog load xdp\_synproxy\_kern.bpf.o /sys/fs/bpf/test\_synproxy type xdp -d
\end{tcolorbox}

\noindent
\textbf{Result}: The verifier \textbf{accepts the program without error}.

\noindent
\textbf{Finding}: No rejection, no warning. The verifier does not validate argument count matching.

\subsection{Stage 4: Xlated Bytecode (bpftool prog dump xlated)}

\vspace{0.3cm}
\noindent
\fcolorbox{black!30}{gray!15}{%
  \begin{minipage}[c]{\dimexpr\textwidth-2\fboxsep-2\fboxrule}
    {\small\ttfamily
    61: (71) r1 = *(u8 *)(r4 +5)\\
    72: (73) *(u8 *)(r2 +6) = r1\\
    73: (18) r1 = map[id:7][0]+0\\
    76: (85) call bpf\_trace\_printk\#-115712
    }
  \end{minipage}%
}
\vspace{0.3cm}

\noindent
\textbf{Finding}: Xlated bytecode is nearly identical to compiled bytecode. The call \textbf{survives verification unchanged}.

\subsection{Stage 5: Runtime Execution (trace\_pipe)}

Trigger the vulnerability:

\begin{tcolorbox}[colback=gray!10, colframe=gray!10, arc=2pt, boxrule=0pt]
\small\ttfamily
python3 src/exploits/5\_06b\_argcomp/exploit\_506b.py
\end{tcolorbox}

\noindent
Kernel trace output:

\vspace{0.3cm}
\noindent
\fcolorbox{black!30}{gray!15}{%
  \begin{minipage}[c][0.8cm][c]{\dimexpr\textwidth-2\fboxsep-2\fboxrule}
    \texttt{<idle>-0 [000] ..s2. 348.738659: bpf\_trace\_printk: [argcomp]: 5.6b - network copy called}
  \end{minipage}%
}
\vspace{0.3cm}

\noindent
\textbf{Finding}: The signal appears repeatedly in kernel traces, \textbf{linking all stages together}. Code executes at runtime.

\section{Exploitation Analysis}

At runtime, an extra argument is passed in R3 (caller setup), but the function consumes only R1 and R2 (callee setup), so R3 is ignored and no immediate memory corruption occurs. The extra argument is ignored by the callee, preventing buffer overflow, pointer dereference, or stack corruption in the current context. However, if the calling context later inspects R3 or other register state, it might find unexpected values. In other contexts, if the mismatched function expected to use the extra arguments, it could access kernel memory via a pointer in R3, use a size parameter in R3 for bounds checking, or perform arithmetic on R3 leading to invalid offsets. In those cases, 5.06b would become \textbf{directly exploitable}.

\section{Artifact Evidence}

All analysis logs are available in: \texttt{src/exploits/5\_06b\_argcomp/logs/}

\begin{itemize}
  \item \texttt{5\_06b\_bytecode\_analysis.txt} — Full llvm-objdump output
  \item \texttt{5\_06b\_verifier\_log.txt} — Verifier output (bpftool prog load -d)
  \item \texttt{5\_06b\_xlated\_dump.txt} — Xlated bytecode (verifier-processed)
\end{itemize}

\noindent
PoC script: \texttt{src/exploits/5\_06b\_argcomp/exploit\_506b.py}
