\chapter{\texorpdfstring{Vulnerability 5.46b: Tainted\\Length (Packet OOB Write)}{Vulnerability 5.46b: Tainted Length (Packet OOB Write)}}

\section{Executive Summary}

Vulnerability 5.46b is a \textbf{confirmed memory corruption vulnerability} where an attacker-controlled \textbf{tainted packet field (TCP Data Offset)} is used directly as a loop counter in a memory write operation, causing a \textbf{buffer overflow within the in-kernel packet buffer}. The corruption writes to adjacent \textbf{packet-local memory} (Ethernet header fields), not kernel data structures.

\begin{tcolorbox}[colback=blue!5!white,colframe=blue!75!black,title=ISO Rule Violated,fonttitle=\bfseries]
ISO-IEC Rule 5.46: ``Do not apply the sizeof operator to a polymorphic object'' (and related: tainted values in sinks)
\end{tcolorbox}

\begin{tcolorbox}[colback=red!5!white,colframe=red!75!black,title=Severity,fonttitle=\bfseries]
\textbf{MEDIUM:} Direct intra-packet memory corruption confirmed, but \textbf{scope-limited to within the packet buffer}. Overwrites adjacent packet fields (MAC addresses, IP/TCP headers), not kernel structures. Extended testing confirmed the verifier \textbf{correctly enforces packet-level bounds} (\texttt{data\_end}); the field-level gap is a \textbf{compiler/language-level issue}. \textbf{No privilege escalation possible}.
\end{tcolorbox}

\subsection{Why This Vulnerability Matters}

\begin{enumerate}[label=\arabic*.]
  \item \textbf{Tainted Packet Field}: TCP Data Offset (doff) is attacker-controlled
  \item \textbf{No Validation}: Value is used directly without bounds checking
  \item \textbf{Fixed-Size Buffer}: Destination MAC address buffer is only 6 bytes
  \item \textbf{Bytecode Limitation}: At the eBPF bytecode level, struct field boundaries (e.g., \texttt{h\_dest[6]}) do not exist; the compiler translates field accesses into flat pointer arithmetic
  \item \textbf{Verifier Behavior}: Correctly enforces packet-level bounds (\texttt{data\_end}), but cannot enforce field-level bounds that are lost during compilation
  \item \textbf{Scope Limitation}: Overflow stays within packet-local memory, not kernel memory
\end{enumerate}

\section{The Vulnerable Code Pattern}

\subsection{Vulnerable Code (Simplified)}

\vspace{0.3cm}
\noindent
\fcolorbox{black!30}{gray!15}{%
  \begin{minipage}[c]{\dimexpr\textwidth-2\fboxsep-2\fboxrule\relax}
    {\small\ttfamily
    /* Fixed-size buffer for MAC destination (6 bytes standard) */\\
    unsigned char h\_dest[6];\\
    \\
    /* Get tainted length from packet TCP header */\\
    \_\_u32 tainted\_length = hdr-\textgreater{}tcp-\textgreater{}doff * 4;\\
    /* doff is 4-15 (4-bit field), so length is 20-60 bytes */\\
    bpf\_printk("[taintsink]: Tainted length \%u", tainted\_length);\\
    \\
    /* Use tainted length in a memory write loop (NO VALIDATION!) */\\
    for (i = 0; i < tainted\_length; i++) \{\\
    \phantom{for (i = 0; i < tainted\_length; i++) \{}h\_dest[i] = /* packet\_data[i] */;\\
    /* BUFFER OVERFLOW if tainted\_length > 6 */\\
    \}
    }
  \end{minipage}%
}
\vspace{0.3cm}

\subsection{Why This Is Vulnerable}

\begin{enumerate}[label=\arabic*.]
  \item \textbf{Fixed-size buffer}: \texttt{h\_dest} is \textbf{only 6 bytes} (standard Ethernet MAC address)
  \item \textbf{Tainted length}: \texttt{doff * 4} is attacker-controlled via TCP Data Offset field
  \item \textbf{Range}: TCP doff is a 4-bit field (values 4-15), so \texttt{doff * 4} = 20-60 bytes
  \item \textbf{No validation}: Code directly uses tainted value as loop counter
  \item \textbf{Overflow}: Writing 60 bytes to a 6-byte buffer = \textbf{54-byte overflow} into adjacent memory
\end{enumerate}

\section{Multi-Stage Analysis: Evidence Chain}

\subsection{Stage 1: Source Code (Patch Location)}

The vulnerability is introduced via a Git patch:
\begin{tcolorbox}[colback=gray!10, colframe=gray!10, arc=2pt, boxrule=0pt]
\small\ttfamily
\texttt{XDPs/xdp\_synproxy/patches/5\_46b\_taintsink/0001-feat-5.46*.patch}
\end{tcolorbox}

\subsection{Stage 2: Bytecode (llvm-objdump)}

Using \texttt{llvm-objdump -S xdp\_synproxy\_kern.o}, the compiled bytecode contains an \textbf{unrolled loop} writing 60 times:

\vspace{0.3cm}
\noindent
\fcolorbox{black!30}{gray!15}{%
  \begin{minipage}[c]{\dimexpr\textwidth-2\fboxsep-2\fboxrule\relax}
    {\small\ttfamily
    ;~ for (i = 0; i < tainted\_length; i++) \{\\
    ;~~ h\_dest[i] = i;\\
    56: (71) r1 = *(u8 *)(r8 +18)    \# Load loop counter i\\
    57: (73) *(u8 *)(r2 -80) = r1   \# Store at offset -80+i\\
    58: (85) call bpf\_trace\_printk  \# Signal: Wrote value at idx\\
    ;~ ... (repeated 60 times for doff=15) ...
    }
  \end{minipage}%
}
\vspace{0.3cm}

\noindent
\textbf{Interpretation}: Bytecode shows 60 store instructions (one for each value in the loop). This confirms the compiler \textbf{did not optimize away} the loop unrolling despite the unsafe length value.

\noindent
\textbf{Finding}: The full 60-iteration loop is preserved in bytecode.

\subsection{Stage 3: Verifier (bpftool prog load)}

Command:
\begin{tcolorbox}[colback=gray!10, colframe=gray!10, arc=2pt, boxrule=0pt]
\small\ttfamily
\texttt{sudo bpftool prog load xdp\_synproxy\_kern.bpf.o /sys/fs/bpf/test\_synproxy type xdp -d}
\end{tcolorbox}

\noindent
\textbf{Result}: The verifier \textbf{accepts the program without error}.

\noindent
\textbf{Finding}: Verifier does not reject the 60-byte write to a 6-byte buffer. Field-level bounds checking missing.

\subsection{Stage 4: Xlated Bytecode (bpftool prog dump xlated)}

The xlated bytecode still contains all 60 store instructions:

\vspace{0.3cm}
\noindent
\fcolorbox{black!30}{gray!15}{%
  \begin{minipage}[c]{\dimexpr\textwidth-2\fboxsep-2\fboxrule\relax}
    {\small\ttfamily
    56: (71) r1 = *(u8 *)(r8 +18)\\
    57: (73) *(u8 *)(r2 -80) = r1\\
    58: (85) call bpf\_trace\_printk\\
    ;~ ... all 60 stores preserved ...
    }
  \end{minipage}%
}
\vspace{0.3cm}

\noindent
\textbf{Finding}: Xlated bytecode shows \textbf{all 60 store instructions are preserved unchanged}. No verifier-inserted bounds checks. Vulnerability survives verification.

\subsection{Stage 5: Runtime Execution (trace\_pipe)}

Trigger the vulnerability with TCP doff=15 (maximum):

\begin{tcolorbox}[colback=gray!10, colframe=gray!10, arc=2pt, boxrule=0pt]
\small\ttfamily
python3 src/exploits/5\_46b\_taintsink/5\_46b\_overflow.py
\end{tcolorbox}

\vspace{0.3cm}
\noindent
\fcolorbox{black!30}{gray!15}{%
  \begin{minipage}[c]{\dimexpr\textwidth-2\fboxsep-2\fboxrule\relax}
    \texttt{<idle>-0 [000] ..s2. 348.737756: bpf\_trace\_printk: [taintsink]: Tainted length 60}\\
    \texttt{<idle>-0 [000] ..s2. 348.738659: bpf\_trace\_printk: [taintsink]: Wrote value at idx 0}\\
    \texttt{<idle>-0 [000] ..s2. 348.739616: bpf\_trace\_printk: [taintsink]: Wrote value at idx 1}\\
    \texttt{<idle>-0 [000] ..s2. 348.740573: bpf\_trace\_printk: [taintsink]: Wrote value at idx 2}\\
    \texttt{<idle>-0 [000] ..s2. 348.741530: bpf\_trace\_printk: [taintsink]: Wrote value at idx 3}\\
    \texttt{<idle>-0 [000] ..s2. 348.757486: bpf\_trace\_printk: [taintsink]: Wrote value at idx 59}\\
    \texttt{<idle>-0 [000] ..s2. 348.758443: bpf\_trace\_printk: [taintsink]: hdr-\textgreater{}eth-\textgreater{}h\_source differs}
  \end{minipage}%
}
\vspace{0.3cm}

\noindent
\textbf{Finding}: The trace confirms:
1. Tainted length is 60 (doff=15)
2. All 60 writes execute (indices 0-59)
3. MAC source field is corrupted (comparison detects difference)

\noindent
\textbf{Links all stages together}. Vulnerability executes at runtime.

\section{Extended Testing: Verifier Boundary Analysis}

To determine the precise boundary of this vulnerability, three additional patch variants were tested. These experiments clarify exactly what the verifier does and does not enforce.

\subsection{Variant A: No Loop Guard}

\noindent
The loop guard was removed entirely, leaving only the tainted \texttt{doff * 4} as the loop bound.

\noindent
\textbf{Result: REJECTED by the verifier.} The verifier detected that a write at offset 74 would reach the end of the allocated packet buffer and returned \texttt{invalid access to packet, offset is outside of the packet}. This confirms the verifier correctly tracks packet pointer bounds.

\subsection{Variant B: Constant Guard (255)}

\noindent
The original tainted guard was replaced with a compile-time constant: \texttt{if (i >= 255) break;}.

\noindent
\textbf{Result: ACCEPTED by the verifier.} The verifier accepts this because the guard is a compile-time constant, giving it a provable upper bound on the loop. Since \texttt{doff * 4} has a maximum of 60 (the \texttt{doff} field is 4 bits), all writes at offsets 0--59 fall within the 74-byte packet buffer. The verifier does not check that \texttt{h\_dest} is only 6 bytes---at the bytecode level, there is no concept of field boundaries within the packet.

\noindent
Runtime output confirmed the OOB writes: the loop wrote to indices 0--59, and the comparison detected \texttt{h\_source} corruption.

\subsection{Variant D: Wider Tainted Source (tcp->window, 16-bit)}

\noindent
\texttt{doff * 4} (4-bit, max 60) was replaced with \texttt{bpf\_ntohs(hdr->tcp->window)} (16-bit, max 65535) as the tainted loop bound, while keeping the constant guard at 255.

\noindent
\textbf{Result: REJECTED by the verifier.} Even with a wider tainted source, the verifier correctly tracked the packet pointer and rejected writes at offset 74 (exactly at the packet boundary), returning \texttt{invalid access to packet, offset is outside of the packet}.

\subsection{Variant Testing Summary}

\begin{table}[h]
\centering
\begin{tabular}{|c|c|c|c|c|}
\hline
\textbf{Variant} & \textbf{Tainted Source} & \textbf{Guard} & \textbf{Max Write} & \textbf{Verifier} \\
\hline
A (no guard) & doff*4 (max 60) & None & 60 & REJECTED \\
B (constant) & doff*4 (max 60) & i>=255 & 60 & ACCEPTED \\
D (wider) & tcp->window (max 65535) & i>=255 & 255 & REJECTED \\
\hline
\end{tabular}
\caption{Variant testing results for vulnerability 5.46b}
\end{table}

\subsection{Conclusions from Variant Testing}

\noindent
The extended testing established three key findings:

\begin{enumerate}[label=\arabic*.]
  \item \textbf{The verifier correctly enforces packet-level bounds.} Writes beyond \texttt{data\_end} are always rejected, regardless of loop guard values. The vulnerability cannot be used to escape the packet buffer and corrupt kernel memory.
  \item \textbf{The verifier does not enforce struct-field-level bounds.} At the eBPF bytecode level, there is no concept of \texttt{h\_dest[6]}. The compiler translates struct field accesses into flat pointer arithmetic (\texttt{data + offset}), and the verifier only checks that this offset stays within \texttt{data}/\texttt{data\_end}.
  \item \textbf{This is a compiler/language-level gap, not a verifier deficiency.} The verifier operates on bytecode where struct field boundaries do not exist. The responsibility for enforcing field-level bounds falls on the compiler or the developer. Neither the C compiler nor the eBPF toolchain inserts field-boundary checks during compilation.
\end{enumerate}

\section{Exploitation Analysis}
\noindent
Intra-packet memory corruption has been \textbf{confirmed}: the verifier accepts the program, the bytecode preserves the 60-byte loop, the xlated bytecode has no bounds guards, runtime executes all 60 writes, and adjacent packet fields are corrupted.

\vspace{0.2cm}
\noindent
Packet memory layout shows that the \texttt{h\_dest} buffer is \textbf{6 bytes} (indices 0-5), \texttt{h\_source} is 6 bytes (indices 6-11), and \texttt{h\_proto} is 2 bytes (indices 12-13). With \textbf{doff=15 (60-byte write)}, bytes 0-5 are written to \texttt{h\_dest} (intended), bytes 6-11 overflow to \texttt{h\_source} (first overflow), bytes 12-13 overflow to \texttt{h\_proto} (second overflow), and bytes 14-59 write into IP/TCP header fields (third overflow).

\vspace{0.2cm}
\noindent
The impact is \textbf{scope-limited}: writes stay within the packet buffer (not kernel stack), the overflow cannot reach kernel memory (confirmed by Variant D testing), the attacker cannot modify task credentials or eBPF context, and while the packet may be corrupted, network processing continues with limited DoS potential.

\vspace{0.2cm}
\noindent
This vulnerability is best understood as a \textbf{compiler/language-level gap}: at the eBPF bytecode level, packet accesses are unstructured pointer arithmetic (\texttt{data + offset}). The C-level struct definitions (\texttt{h\_dest[6]}, \texttt{h\_source[6]}, etc.) are an abstraction that exists only at the source code level and is lost during compilation. The verifier correctly enforces the packet-level boundary (\texttt{data\_end}), but has no mechanism to enforce field-level boundaries that do not exist in the bytecode.

\vspace{0.2cm}
\noindent
The vulnerability is \textbf{exploitable within packet-buffer scope}: intra-packet memory corruption is confirmed, but the eBPF sandbox and verifier packet bounds prevent kernel-level compromise.


\section{Artifact Evidence}

All analysis logs available in: \texttt{src/exploits/5\_46b\_taintsink/logs/}

\begin{itemize}
  \item \texttt{5\_46b\_bytecode\_analysis.txt} — Full llvm-objdump output (all 60 stores)
  \item \texttt{5\_46b\_verifier\_log.txt} — Verifier output (bpftool prog load -d)
  \item \texttt{5\_46b\_xlated\_dump.txt} — Xlated bytecode with all 60 store instructions
\end{itemize}

\noindent
PoC script: \texttt{src/exploits/5\_46b\_taintsink/5\_46b\_overflow.py}

